### 2014-10-24 thesis-writing/methodeval.tex

\section{Methodology and Evaluation}
I am on the process of implementing the garbage collector in the OpenJDK project. The plan to evaulate the garbage collector algorithm for the shared memory is to implement in the HotSpot Virtual Machine in OpenJDK project and measure the performances in terms of various parameters to prove the hyphothesis is right.