### 2014-10-24 thesis-writing/timeline.tex

\section{Timeline}
\begin{table}
\caption{Proposed Timeline for obtaining degree.}
\centering
\begin{minipage}[t]{.7\linewidth}
\color{gray}
\rule{\linewidth}{1pt}
\ytl{Dec, 2014}{Finish the implementation of shared memory GC in a virtual machine and obtain the benchmark results.}
\ytl{Feb, 2015}{Publish the performance of the GC in Conference/Journal}
\ytl{May, 2015}{Design a solution for the distributed GC algorithm}
\ytl{Aug, 2015}{Complete the implemantation of distributed GC in a runtime system}
\ytl{Dec, 2015}{Publish the results in Conference/Journal}
\ytl{Mar, 2016}{Complete writing thesis and defend}
\ytl{May, 2016}{Graduate!}
\bigskip
\rule{\linewidth}{1pt}%
\end{minipage}%
\end{table}