### 2014-10-31 thesis-writing/methodeval.tex
last minute changes

\section{Methodology and Evaluation}
I am on the process of implementing the garbage collector in the OpenJDK project. The plan to evaluate the garbage collector  is to implement in the HotSpot Virtual Machine in OpenJDK project and measure the performances in terms of various parameters to prove the hypothesis is right. It is necessary to note that the algorithm is not designed for the environment where it is in the process of implementing. The parameter that is under consideration are throughput, pause time,  and scalability. While there is no formal distributed garbage collector framework to test the distributed garbage collection algorithm, Proactive ~\cite{activeon} and X10 ~\cite{x10} are two venues to implement the garbage collectors and measure performance. To my knowledge, all works published in this area has never used any benchmark to provide enough empirical evidences to prove that their distributed collector performed better than alternatives. So this research focuses on using custom built benchmarks to test the efficiency of the algorithm.