### 2016-05-19 finaletd/abstract.tex
distributed paper added

\begin{abstract}

We propose a scalable, cycle-collecting, decentralized, reference counting
garbage collector with partial tracing.  The algorithm is based on the
Brownbridge system but uses four different types of references to label edges
in the reference graph. Memory usage is $O(\log n)$ bits per node, where $n$ is the
number of nodes in the graph.  The algorithm assumes an asynchronous network
model with a reliable FIFO channel. It collects garbage in
$O(E)$ time, where $E$ is the number of edges in the induced subgraph of the reference graph. The
algorithm uses termination detection to manage the
distributed computation, a unique identifier to break the symmetry among
multiple collectors, and a transaction-based approach 
when multiple collectors conflict. Unlike existing
algorithms, ours is not centralized, does not require barriers,
does not require
migration of nodes, does not require back-pointers on every edge,
and is stable against concurrent mutation. %, even with multiple collection
%operations overlapping the induced subgraph.

%\keywords{Distributed Garbage Collection, Distributed Termination Detection,
%Strong Weak Phantom, Reference Count}
\end{abstract}