### 2016-05-19 finaletd/model.tex

\section{Model and Preliminaries}
\label{model}
\paragraph{Distributed System:}
We consider a distributed system of nodes
%A distributed algorithm is one where
%self-stabilization problem where nodes
%the dead nodes identify that they are dead and
%delete themselves. Identifying a dead node by itself is not trivial in most
%cases. Each node can be assumed to
where each node operates independently and communicates with other nodes through
message passing. The underlying topology of the system is assumed to be arbitrary but connected. Each node has a queue to receive messages, and in response to a message a node can only read and write its own state and send additional messages. %A node can only send a
%message to its out-neighbors and parent of the node. Parent, a chosen node in
%$\Gamma_{in}$ from communication happened by our algorithm, .
We further assume that the nodes can communicate over
a reliable FIFO channel and that messages
are never lost, duplicated, or delivered out of order.
These properties make our system compatible with the $\mathcal{CONGEST}$ asynchronous network model with no failures ~\cite{congest}.
%Each site has memory where the objects reside. \todo{need precise definition}

\paragraph{Basic Reference Graph:}
We model the relationship among various objects and pointers in memory through a directed graph $G = (V, E)$, which we call a {\it reference graph}.
The graph $G$ has a special node $R$, which we call the {\it root}. Node $R$
represents
global and stack pointer variables, and
thus does not have any incoming edges.
Each node in $G$ is assumed to contain a unique ID. % to distinguish it within
%$G$. (For example, when each node of $G$ maps to a unique site in $H$, then ID of the site in $H$ can be used for the ID of the node in $G$.)
All adjacent nodes to a given node in $G$ are called $neighbors$, and denoted by $\Gamma$. The
$\inneighbors$ of a node $x\in G$ can be defined as the set of nodes whose outgoing edges
are incident on $x$, represented by $\Gamma_{in}(x)$. The $\outneighbors$ of $x$ can be defined as the set
of nodes whose incoming edges originate on $x$, represented by $\Gamma_{out}(x)$.
Note that each node $x\in G$ does not know $\Gamma_{in}(x)$ at any point in time.
%This constraint
%reduces memory complexity, thereby making
%the algorithm more practical and scalable. \todo{remove preceding sentence?}

\paragraph{Distributed Garbage Collection Problem:}
All nodes in $G$ can be classified as either {\em live} (i.e., not garbage) or {\em dead} (i.e., garbage) based on a property called
$reachability$. Live and dead nodes can be defined as below:

%$Live(x)= \exists y \in V \mid Reachable(y, x)$

%$Reachable(y, x) = x \in R \vee (Live(y) \wedge x \in \Gamma_{out}(y))$

$Reachable(y,x) = x \in \Gamma_{\rm out}(y) \lor (x \in \Gamma_{\rm out}(z) \mid Reachable(y,z))$

$Live(x) = Reachable(R,x)$

$Dead(x) = \neg Live(x)$

%From above definition, it is clearly that all root nodes are always reachable / live.
We allow the live portion
of $G$, denoted as $G'$, to be mutated while the algorithm is running, and we refer
to the source of these mutations as the {\em Adversary.}
The Adversary can create nodes and attach them to $G'$, create new edges between existing nodes of $G'$, or delete edges from $G'$. Moreover, the Adversary can perform multiple events (creation and deletion of edges)
simultaneously. The Adversary, however, can never mutate the dead portion of the graph $G'' = G\backslash G'$.
%Creation event represents creating directed
%edges. An edge can be only created between two live nodes in $G$. To create a
%node, an edge is created between a live node and non-existing node. During link
%creation, if the node is not available, the requested node is created and a
%directed edge is incident on newly created node. So the newly
%created node is always reachable. Deletion happens only for edges. Adversary can
%delete outgoing edge of a node $x$ only if $x$ is live . So $A$ cannot perform
%any mutation on dead nodes. These constraints create the following principles /
%axioms in the model.

\begin{axiom}[Immutable Dead Node]
 The Adversary cannot mutate a dead node.
 \label{ax:immut}
\end{axiom}

\begin{axiom}[Node Creation]
 All nodes are live when they are created by the Adversary.
\end{axiom}

From Axiom \ref{ax:immut}, it follows that a node that becomes dead will never
become live again.
%This property allows the problem to be solved in
%finite time. \todo{remove preceding sentence}
%The problem of distributed garbage collection can be modeled as
%detecting all dead nodes in dynamic directed graph $G$.

Each node experiencing deletion of an incoming edge has to determine whether it is still
live. If a node detects that it is dead then it must delete itself from the
graph $G$. %The problem of garbage collection is the process of
%identifying dead nodes in $G$ and deleting them.

\begin{definition}[Distributed Garbage Collection Problem]
%The problem of garbage collection is the process of
Identify the dead nodes in the reference graph $G$ and delete them.
\end{definition}

%We consider a distributed system where
%%A distributed algorithm is one where
%%self-stabilization problem where nodes
%%the dead nodes identify that they are dead and
%%delete themselves. Identifying a dead node by itself is not trivial in most
%%cases. Each node can be assumed to
%each node communicates with other nodes through
%messages. Each node has a queue to receive messages, and in response to a message a node can only read and write its own state and send additional messages. %A node can only send a
%%message to its out-neighbors and parent of the node. Parent, a chosen node in
%%$\Gamma_{in}$ from communication happened by our algorithm, .
%We further assume that the algorithm communicates over
%a reliable FIFO channel and that messages
%are never lost, duplicated, or delivered out of order.
%These properties make this algorithm compatible with the $\mathcal{CONGEST}$~\cite{congest} asynchronous network model with no failures.  \todo{talk also about fair execution here which is currently missing}
%%The $\mathcal{CONGEST}$ model adds the additional requirement that %messages are limited in size to $O (\log N)$. This limit is %already reached if unique object ID's are present in a message, as %they are of size $O (\log N)$.

%\subsection{Preliminaries}

\begin{figure}
\centering
%\begin{center}
\scalebox{0.6}[0.6]{\input{completeabstract}}
\caption{We depict all the ways in which an initiator node, $A$, can
be connected to the graph. Circles 
represent sets of nodes.
Dotted lines represent one or more non-strong paths. Solid lines represent one or more
strong paths.
A T-shaped end-point indicates the root, R.
If C${}^\prime$, D${}^\prime$, E${}^\prime$ and F are empty sets, A is garbage, otherwise 
it is not. }
\label{fig:completeabstract}
%\end{center}
\end{figure}

\begin{comment}
\begin{figure}[h!]
\centering
\begin{subfigure}[t]{0.45\linewidth}
%\begin{center}
\scalebox{0.6}[0.6]{\input{completeabstract}}
\caption{We depict all the ways in which an initiator node, $A$, can
be connected to the graph. Circles 
represent sets of nodes.
Dotted lines represent one or more non-strong paths. Solid lines represent one or more
strong paths.
A T-shaped end-point indicates the root, R.
If C', D', E' and F are empty sets, A is garbage, otherwise 
it is not. }
\label{fig:completeabstract}
%\end{center}
\end{subfigure}
%\hfill
\begin{subfigure}[t]{0.45\linewidth}
%\begin{center}
\scalebox{0.6}[0.6]{\input{figs/state2}}
\caption{The above figure depicts the state/phase transitions performed by initiator in the algorithm.
}
\label{fig:state}
%\end{center}

\end{subfigure}
\caption{}
\end{figure}
\end{comment}
\begin{comment}

\begin{figure}
\begin{center}
\scalebox{0.6}[0.6]{\input{completeabstract}}
\caption{An illustration of all the ways in which an initiator node, $A$, can
be connected to the graph before a collection process begins. The circles in the
graph represent sets of nodes.
Dotted lines represent one or more non-strong paths, and solid lines represent one or more
strong paths (but non-strong paths may exist alongside them).
A T-shaped end-point indicates the root, $R$.
Other nodes and connections between nodes may exist,
but here we map only paths and nodes related to $A$ and $R$.
Thus, if $C'$, $D'$, $E'$ and $F$ are empty sets, $A$ is garbage, otherwise
it is not.}
\label{fig:completeabstract}
\end{center}
\vspace{-9mm}
\end{figure}
\end{comment}
%\paragraph{Modified Reference Graph $G$ using the Brownbridge Method:}
\paragraph{Classification of Edges in $G$:}
%Instead of using the reference graph $G$ defined above, we use the Brownbridge method \cite{Brownbridge1985} of classifying edges of $G$.
We classify the edges in $G$ as $weak$ or $strong$ according to the Brownbridge method~\cite{Brownbridge1985}.
This classification does not need preprocessing of $G$ and can be done directly at the time of creation of $G$.
%In the Brownbridge method, edges of $G$ are labeled as either $strong$ or $weak$. %Each node has two reference counts, one for each
%incoming edge type.
%The reason for the two types of edges are due to
%inability to detect cyclic structure by single edge type.
%Brownbridge idea is
%to avoid cycle of strong edges in the graph.
Chains of strong edges connect the root $R$ to all the nodes $G$ and contain
no cycles. The weak edges
are required for all cycle-creating edges in $G$. The advantage to the Brownbridge
method is that it permits less frequent
tracing of the outgoing edges. Without a distinction between strong and weak edges,
tracing is required after every deletion that does not bring the reference count to zero.
In the Brownbridge method, if a strong edge remains after the deletion of any edge,
the node is live and no tracing
is necessary.
%Our algorithm takes graph $G$ as this modified reference graph.
Fig.~\ref{fig:completeabstract} illustrates an example reference graph $G$ with strong (solid) and weak (dotted) edges.

\paragraph{Weight-based Edge Classification for the Reference Graph $G$:} We assign {\em weights} to the nodes as a means of classifying the edges in $G$.
Each node of $G$ maintains two attributes %related to the weight component of the algorithm.
$weight$ and $\maxweight$, as well as counts of strong and weak incoming edges. %The weight is an integer in each node that helps to
%label the edges in the graph.
The weight of the root, $R$, is zero at all times. For
an edge to
be strong, the weight of the source must be less than the weight of the
target.
The
$\maxweight$ of a node is defined as the maximum weight of the node's
$\inneighbors$. Messages sent by nodes contain
the $weight$ of the sender so that the $\outneighbors$ can keep track of their
strong count, weak count, and $\maxweight.$
When a
node is created, its initial weight is given by adding one to the weight of the
first node to reference it.
%edge to the node is a strong edge. An edge is considered strong only when the
%weight of the source node is less than weight of the target node. So a strong
%edge $e$ from $x$ to $y$ means $x.weight$ $<$ $y.weight$. Node creation followed by
%edge created to that node updates the $\maxweight$ of the node. If the node is
%just created, the weight is updated to the first increment of max weight. This
%satisfies the requirements to make the edge into strong edge.

%The advantage of
%weight based labeling is, along with atomic edge labeling, our algorithm
%requires
%the edge label to be changed
%without the source of the edge interacting with the conversion process. In this
%approach, a
%target node of an edge can change the label from strong to weak or vice-versa by
%adjusting
%its weight. When a node creates an edge, the source of the edge does not
%wait for the target to notify the label of the edge because the label of the
%edge is based on
%the weight of the source which is already sent in the message to create that
%edge.
%--% \begin{figure}
%--% \begin{center}
%--% \scalebox{0.6}[0.6]{\input{figs/fig2}}
%--% \caption{Examples of labeling edges when the edge is created}
%--% \label{fig:brownbridge}
%--% \end{center}
%--% \end{figure}

%\begin{figure}[h]
%\caption{Example of labeling edges when the edge is created using brownbridge
%technique}
%\centering
%\includegraphics[width=\textwidth]{figs/strongweak}
%\label{fig:brownbridge}
%\end{figure}

%Note that having a weak incoming edge does not necessarily mean the
%node is part of cycle.
% as depicted in the \ref{fig:brownbridge}.
%In figure \ref{fig:brownbridge}, there are two examples shown explaining
%our technique to name edges. %$N_i$ can be considered to be created
%before $N_{i-1}$.
%In the image on the left, a weak edge is labeled to mark
%the cycle closing edge. On the right side, however, an edge is marked weak although
%node $N_4$, which is not part of any cycle.

%\todo{remove?}
\begin{lemma}[Strong Cycle Invariant]
No cycle of strong edges will be formed by weight-based edge classification.
\end{lemma}

\begin{proof}
By construction, for any node $y$ which is reachable from a node $x$ by
strong edges, $y$ must have
a larger weight than $x$. Therefore, if $x$ is reachable from itself through a cycle,
and if that cycle is strong, $x$ must have a greater
weight than itself, which is impossible.
%To create a cycle of strong edges, the weight of source of cycle closing edge
%must be lesser than the weight of the target node. Let's assume the source of
%cycle closing edge for the graph is $z$ and target of the cycle closing edge
%is $a$. As we mentioned above, the cycle closing edge can be strong only if
%the $z.weigh$ $<$ $a.weight$. For this to be a cycle closing edge, there must
%be at least an edge from $a$ to series of nodes ultimately reaching $z$ or
%$a$ contains a edge to $z$ directly. In the later case, it is clear that the
%edge from $a$ to $z$ will be weak since the weights are not satisfying the
%condition to label them as strong. In the former case, for all the edges
%between $a$ to $z$ to be strong, all the weight of nodes in the path from
%$a$ to $z$ must be in ascending order. This contradicts our assumption that
%$a.weight$ $>$ $z.weight$. So a cycle of strong edges cannot be formed in weightbased edge labeling.
\end{proof}

\paragraph{Terminology:}
A $path$ is an ordered set of edges such that the destination
of each edge is the origin of the next.
A path is called a $strong$ path if it consists
exclusively of strong edges, otherwise it is called
$\nonstrong$ path. We say two nodes are $related$ if
a path exists between them.
A node, $x$, is a member of the $dependent$ $set$ of A if all strong paths leading to $x$ originate on $A$
(i.e. $B$, $B'$, $C$, and $C'$ in Fig.~\ref{fig:completeabstract}).

Consider a node, A, that has just lost its last strong edge. We model this situation
with an induced subgraph with $A$ removed (i.e. $G-A$), then the $purely$ $dependent$ $set$ of A
consists of the nodes that have no path from $R$
(i.e. $B$, and $B'$  in Fig.~\ref{fig:completeabstract}).
A node, $y$, is a member of the $supporting$ $set$ of A if
there is a path from $R$ to $y$, and from $y$ to $A$ (i.e.
$C'$, $D'$, $E'$, and $F$ in Fig.~\ref{fig:completeabstract} ).
A node, $z$, is a member of the $independent$ $set$ of A if $z$ is related to $A$ and
there is at least one strong path from $R$ to $z$
(i.e.
$D$, $D'$, $E$, $E'$, and $F$ in Fig.~\ref{fig:completeabstract}).
A node, $x$, is a member of the $build$ $set$ of A if it is both a member of the supporting set and
the independent set.
Alternatively, A node, $x$ is said to be in the $build$ $set$ if a strong path from $R$ to $x$ exists,
but also a path from $x$ to $A$ exists (i.e. $D'$, $E'$ and $F$ in Fig.~\ref{fig:completeabstract}).
A node, $x$, is a member of the $recovery$ $set$ of A if it is both a member of the supporting set
and the dependent set
(i.e. $C'$ in Fig.~\ref{fig:completeabstract}).
A node, $x$, is said to be in the $\mathit{affected\ set}$ if there exists a path, $p$, from $A$ to $x$, such that no
node found along $p$ (other than $x$ itself) has a strong path from $R$ (i.e.
everything except $F$ in Fig.~\ref{fig:completeabstract}).