### 2016-05-20 finaletd/brownbridge.tex
Thesis sections and chapters added with some clear sections to fill

\section{Cyclic Graph}
\section{Cycle detection using Strong-Weak heuristics}
\section{Brownbridge Garbage Collector}
\section{Pitfalls of Brownbridge and successor's technique}