### 2016-06-02 finaletd/mainconclusion.tex
conclusion

In the multi-core and distributed systems era, the necessity of a complete high performance garbage collector is necessary. The garbage collection algorithms discussed in the thesis is designed for models that mimic the reality in shared memory and distributed memory systems. So these garbage collectors does not require any changes to be implemented in any programming language environment. The assumptions considered in both the environments to solve holds true in existing systems in the environments. The thesis contributions can be summarized as below:
\begin{itemize}
		\item Presented a novel cyclic hybrid reference counting and tracig technique garbage collector for general graph.
		\item Designed first known concurrent multi-collector shared memory garbage collector.
		\item Flawed Brownbridge garbage collector algorithm is corrected and extended by our garbage collectors.
		\item Reduced the significant number of traversal required to detect garbage in dynamic concurrent garbage collection environment when compared to state of the art reference counted concurrent garage collectors.
		\item Introduced first known distributed garbage collection with no global synchronization.
		\item Introduced first known locality-based distributed garbage collection.
		\item Designed highly scalable garbage collectors for both enviornments.
		\item Introduced low memory cost, low communication cost weight based approach to convert the reference graph into directed acyclic graph and thereby detec cycle faster.
		\item Both garbage collectors operate in linear time.
\end{itemize}